Vestibular Contributions to Spatial Perspective Transformations: From Sensory Inference To Imagined Self-Motion

### 2017-01-25 manuscript.Rmd
fixed word conversion problem

lang              : "english"
class             : "man"
# output            : papaja::apa6_word
output:
     # papaja::apa6_word
     papaja::apa6_pdf:
        pandoc_args:
        fig_caption: true
---

@@

knitr::opts_chunk$set(cache = TRUE,
                      fig.width = 12, fig.height = 8,
                      dpi = 600, fig.path = "../generated-figures/",
                      dev=c('pdf', 'png', 'cairo_ps'))
```

@@

![Perspective taking and imagined self-motion involve spatial transformations of a representation of the self. (a) If a speaker's goal is to take the perspective of an addressee, he must first solve the inverse problem of determining the appropriate action, and then run a forwards simulation in order to perform a translation and rotation of his own egocentric reference frame. These operations can be viewed as imagined movements.(b) Focussing on imagined rotation around the vertical axis allows us to simplify the problem, and highlights the fact that perspective taking and mental imagery involve common computations.](diagrams/fig-1-common-process.pdf){#fig:spt-schematic width=75%}

@@

```{r head-turn-data, echo=FALSE, fig.cap= "Sensory measurements obtained during a short leftward head turn. The goal is to infer the latent angular velocity and position, based on the noisy measurements. \\label{fig:head-turn}", message=FALSE, warning=FALSE, dpi=100}
dt <- 0.1
motiondata <-  generate_data(T = 2, frequency = 0.5,
                             amplitude = 20,
                             sensor_sd = 2.0)
# plot_trajectories(motiondata = motiondata)

motiondata_hidden <- motiondata %>%
    gather(type, value = value, -time) %>%
    filter(type != "observations") %>%
    mutate(type = factor(type,
                         levels = c("position",
                                    "velocity",
                                    "acceleration")))

p1 <- motiondata_hidden %>%
        ggplot(aes(x = time, y = value)) +
            geom_line(aes(linetype = type), size = 1.4) +
            geom_point(data = motiondata,
                       aes(x = time, y = observations),
                       shape = 21, colour = "white",
                       fill = "white", size = 6, stroke = 2) +
            geom_point(data = motiondata,
                       aes(x = time, y = observations),
                       shape = 21, colour = "black",
                       fill = "white", size = 4, stroke = 2) +
            guides(linetype = guide_legend(keywidth = 2.5, keyheight = 1)) +
            xlab("Time (sec)") +
            ylab("Angular position (deg)") +
            theme_apa() +
            theme(legend.title=element_blank())
# p1
plot_trajectories(motiondata = motiondata)
```

@@

```{r graphical-model-inference, fig.cap="Graphical model for inferring the 2-dimensional state $x$, consisting of angular velocity and position, based on noisy sensory measurements $y$.", echo=FALSE}
knitr::include_graphics('diagrams/dynamic-belief-network-inference-1.pdf')
```

@@

```{r particle-filter-explained, fig.cap="Particle filter computations performed at every time step. The cloud of particles (grey circles) represents the current belief state. Initially, a prediction is made for each particle by sampling from the proposal distribution, which is determined by the agent’s kinematic model. This is equivalent to sampling from the prior distribution. Then, during the updating step, an observation  is obtained (shown as a black diamond). Each particle is assigned an importance weight, according to how well it explains the observation. Each particle's diameter reflects its importance weight. Subsequently, a new set of particles is formed by weighted multinomial resampling, with each particle's probability of being chosen proportional to its importance weight. The particle cloud's mean and standard deviation are shown as light grey circles and lines. This resulting cloud of particles represents the posterior distribution. This posterior subsequently forms the prior distribution for the next time step.", echo=FALSE}
knitr::include_graphics('diagrams/figure-particle-filter-explained.pdf')
```

@@

```{r eval=FALSE, include=FALSE}
plot_pf_estimates(object = passive, predict = FALSE)
```

@@

$$\theta_t = \theta_{t-1} + \omega_{t-1} \Delta t + \alpha_{t-1} \frac{1}{2}\Delta t^2$$

$$\omega_{t} = \omega_{t-1} + \alpha_{t-1} \Delta t$$
The effect that this has in terms of the inference algorithm is that the proposal distribution $P(x_t | y_{1:t})$, from which samples are drawn from during the 'predict' step, is improved. By including the acceleration in the latent state, the the changes in velocity are modelled. Of course, the acceleration itself is not modelled, and the changes in acceleration (jerk) are thus treated as random, resuling in a random jerk model.

@@

```{r shaking-head, eval=FALSE, fig.cap="Particle filter applied to...", message=FALSE, warning=FALSE, cache=TRUE, include=FALSE}
dt <- 0.1
head_shaking_data <- generate_data(T = 2, frequency = 2.0,
                             amplitude = 20,
                             sensor_sd = 2.0)

@@

```{r two-external-forces, eval=FALSE, fig.cap="Particle filter applied to...", message=FALSE, warning=FALSE, cache=TRUE, include=FALSE}
dt <- 0.1
ext_data_1 <- generate_data(T = 4, frequency = 0.7,
                             amplitude = 20,
                             sensor_sd = 2.0)
ext_data_2 <- generate_data(T = 4, frequency = 0.5,
                             amplitude = -5,
                             sensor_sd = 2.0)
superposition_data <- ext_data_1 + ext_data_2

@@

```{r graphical-model-control, fig.cap="Graphical model extended with hierarchical model for self-initiated movements.", echo=FALSE}
knitr::include_graphics('diagrams/dynamic-belief-network-inference-2.pdf')
```

@@

THe brain implements multiple priors:

@@

control input -> infer the consequences of one's own actions.

@@

- place the emerging field of vestibular cognition with a framewokk for dynamic probabilistic inference and draw links to established models of sensory processing

>>>>

lang              : "english"
class             : "man"
output:
     # papaja::apa6_word:
     papaja::apa6_pdf:
        pandoc_args:
            fig_caption: true
---

@@

knitr::opts_chunk$set(cache = TRUE,
                      fig.width = 12, fig.height = 8,
                      dpi = 600, fig.path = "generated-figures/",
                      dev=c('pdf', 'png', 'tiff', 'cairo_ps'))
```

@@

<!-- ![Perspective taking and imagined self-motion involve spatial transformations of a representation of the self. (a) If a speaker's goal is to take the perspective of an addressee, he must first solve the inverse problem of determining the appropriate action, and then run a forwards simulation in order to perform a translation and rotation of his own egocentric reference frame. These operations can be viewed as imagined movements.(b) Focussing on imagined rotation around the vertical axis allows us to simplify the problem, and highlights the fact that perspective taking and mental imagery involve common computations.](diagrams/fig-1-common-process.tiff){#fig:spt-schematic width=75%} -->

```{r spt-schematic, fig.cap="Perspective taking and imagined self-motion involve spatial transformations of a representation of the self. (a) If a speaker's goal is to take the perspective of an addressee, he must first solve the inverse problem of determining the appropriate action, and then run a forwards simulation in order to perform a translation and rotation of his own egocentric reference frame. These operations can be viewed as imagined movements.(b) Focussing on imagined rotation around the vertical axis allows us to simplify the problem, and highlights the fact that perspective taking and mental imagery involve common computations.", echo=FALSE}
knitr::include_graphics('diagrams/fig-1-common-process.tiff')
```

@@

```{r head-turn-data, echo=FALSE, fig.cap= "Sensory measurements obtained during a short leftward head turn. The goal is to infer the latent angular velocity and position, based on the noisy measurements. \\label{fig:head-turn}", message=FALSE, warning=FALSE, dpi=100}
dt <- 0.1
motiondata <-  generate_data(T = 2, frequency = 0.5,
                             amplitude = 20,
                             sensor_sd = 2.0)
plot_trajectories(motiondata = motiondata)

# motiondata_hidden <- motiondata %>%
#     gather(type, value = value, -time) %>%
#     filter(type != "observations") %>%
#     mutate(type = factor(type,
#                          levels = c("position",
#                                     "velocity",
#                                     "acceleration")))
# 
# p1 <- motiondata_hidden %>%
#         ggplot(aes(x = time, y = value)) +
#             geom_line(aes(linetype = type), size = 1.4) +
#             geom_point(data = motiondata,
#                        aes(x = time, y = observations),
#                        shape = 21, colour = "white",
#                        fill = "white", size = 6, stroke = 2) +
#             geom_point(data = motiondata,
#                        aes(x = time, y = observations),
#                        shape = 21, colour = "black",
#                        fill = "white", size = 4, stroke = 2) +
#             guides(linetype = guide_legend(keywidth = 2.5, keyheight = 1)) +
#             xlab("Time (sec)") +
#             ylab("Angular position (deg)") +
#             theme_apa() +
#             theme(legend.title=element_blank())
# p1
```

@@

```{r graphical-model-inference, fig.cap="Graphical model for inferring the 2-dimensional state $x$, consisting of angular velocity and position, based on noisy sensory measurements $y$.", echo=FALSE}
knitr::include_graphics('diagrams/dynamic-belief-network-inference-1.tiff')
```

@@

```{r particle-filter-explained, fig.cap="Particle filter computations performed at every time step. The cloud of particles (grey circles) represents the current belief state. Initially, a prediction is made for each particle by sampling from the proposal distribution, which is determined by the agent’s kinematic model. This is equivalent to sampling from the prior distribution. Then, during the updating step, an observation  is obtained (shown as a black diamond). Each particle is assigned an importance weight, according to how well it explains the observation. Each particle's diameter reflects its importance weight. Subsequently, a new set of particles is formed by weighted multinomial resampling, with each particle's probability of being chosen proportional to its importance weight. The particle cloud's mean and standard deviation are shown as light grey circles and lines. This resulting cloud of particles represents the posterior distribution. This posterior subsequently forms the prior distribution for the next time step.", echo=FALSE}
knitr::include_graphics('diagrams/figure-particle-filter-explained.tiff')
```

@@

```{r eval=FALSE, include=FALSE}
# plot_pf_estimates(object = passive, predict = FALSE)
```

@@

$$ \theta_t = \theta_{t-1} + \omega_{t-1} \Delta t + \alpha_{t-1} \frac{1}{2}\Delta t^2 $$

$$ \omega_{t} = \omega_{t-1} + \alpha_{t-1} \Delta t $$

The effect that this has in terms of the inference algorithm is that the proposal distribution $P(x_t | y_{1:t})$, from which samples are drawn from during the 'predict' step, is improved. By including the acceleration in the latent state, the the changes in velocity are modelled. Of course, the acceleration itself is not modelled, and the changes in acceleration (jerk) are thus treated as random, resuling in a random jerk model.

@@

```{r shaking-head, eval=FALSE, fig.cap="Particle filter applied to ", message=FALSE, warning=FALSE, cache=TRUE, include=FALSE}
dt <- 0.1
head_shaking_data <- generate_data(T = 2, frequency = 2.0,
                             amplitude = 20,
                             sensor_sd = 2.0)

@@

```{r two-external-forces, eval=FALSE, fig.cap="Particle filter applied to ", message=FALSE, warning=FALSE, cache=TRUE, include=FALSE}
dt <- 0.1
ext_data_1 <- generate_data(T = 4, frequency = 0.7,
                             amplitude = 20,
                             sensor_sd = 2.0)
ext_data_2 <- generate_data(T = 4, frequency = 0.5,
                             amplitude = -5,
                             sensor_sd = 2.0)
superposition_data <- ext_data_1 + ext_data_2

@@

```{r graphical-model-control, fig.cap="Graphical model extended with hierarchical model for self-initiated movements.", echo=FALSE}
knitr::include_graphics('diagrams/dynamic-belief-network-inference-2.tiff')
```

@@

The brain implements multiple priors:

@@

control input allows inferring the consequences of one's own actions.

@@

- place the emerging field of vestibular cognition with a framework for dynamic probabilistic inference and draw links to established models of sensory processing
